Combinatorial Proofs and Decomposition Theorems for First-order Logic

### 2021-01-10 main.tex
Complete the negation part

In this paper we solve this issue by moving to a deep inference
system. More precisely, we introduce a new proof system, called
$\FOKS$, for first-order logic, that (i) reflects every combinatorial
proof, i.e., there is a surjective mapping from proofs in $\FOKS$ to
combinatorial proofs, that (ii) allows to provide simpler proofs of
soundness and completeness of combinatorial proofs, and (iii) admits
new decomposition theorems establishing a precise correspondence
between certain syntactic inference rules and certain combinatiorial
notions.

@@

We start from a countable set $\VAR=\set{x,y,z,\ldots}$ of variables,
a countable set $\FUN=\set{f,g,\ldots}$ of function symbols, and a
counrable set $\PRED=\set{p,q,\ldots}$ of predicate symbols. Each function symbol and
each predicate symbol has a finite arity. The grammars below generate
the set $\TERM$ of \bfit{terms}, denoted by $s,t,u,\ldots$, the set
$\ATOM$ of \bfit{atoms}, denoted by $a,b,c,\ldots$, and the set
$\FORM$ of \bfit{formulas}, denoted by $A,B,C,\ldots$:
\begin{equation*}
  \begin{array}{r@{~}l}
    t &\coloneqq~ x \mid f(t_1,\dots,t_n)
    \\[.5ex]
    a &\coloneqq~ \ttt\mid\fff\mid p(t_1,\dots,t_n) \mid \dual p(t_1,\dots,t_n)
    \\[.5ex]
    A &\coloneqq~ a \mid A\vlan A \mid A\vlor A \mid \exists x.A \mid \forall x.A
\end{array}
\end{equation*}
where the arity of $t$ and $p$ is $n$. Note, that in this paper we
consider the truth constants $\ttt$ (\emph{true}) and $\fff$
(\emph{false}) as atoms, and we consider all formulas in negation
normalform. The \bfit{negation} $(\dual\cdot)$ is defined for all
atoms and formulas via the De Morgan laws as follows:
\begin{equation*}
  \dual\ttt =\fff
  \quad
  \dual\fff = \ttt
  \quad
  \dual{p(t_1 \ldots t_n)} = \dual{p}(t_1 \ldots t_n)
  \quad
  \dual{\dual p(t_1 \ldots t_n)} = p(t_1 \ldots t_n)
\\
  \dual{\dual a} = a
\end{equation*}

@@

 A formula is
\bfit{rectified} if all bound variables are distinct from one another and from
all free variables. As every formula has a logically equivalent
rectified form, we will most of the results in this paper to rectified formulas.

>>>>

In this paper we solve this issue by moving to a deep inference
system. More precisely, we introduce a new proof system, called
$\FOKS$, for first-order logic, that (i) reflects every combinatorial
proof, i.e., there is a surjective mapping from proofs in $\FOKS$ to
combinatorial proofs, that (ii) allows to provide simpler proofs of
soundness and completeness of combinatorial proofs, and (iii) admits
new decomposition theorems establishing a precise correspondence
between certain syntactic inference rules and certain combinatorial
notions.

@@

We start from a countable set $\VAR=\set{x,y,z,\ldots}$ of variables,
a countable set $\FUN=\set{f,g,\ldots}$ of function symbols, and a
counrable set $\PRED=\set{p,q,\ldots}$ of predicate symbols. Each function symbol and
each predicate symbol has a finite arity. The grammars below generate
the set $\TERM$ of \bfit{terms}, denoted by $s,t,u,\ldots$, the set
$\ATOM$ of \bfit{atoms}, denoted by $a,b,c,\ldots$, and the set
$\FORM$ of \bfit{formulas}, denoted by $A,B,C,\ldots$:
\begin{equation*}
  \begin{array}{r@{~}l}
    t &\coloneqq~ x \mid f(t_1,\dots,t_n)
    \\[.5ex]
    a &\coloneqq~ \ttt\mid\fff\mid p(t_1,\dots,t_n) \mid \dual p(t_1,\dots,t_n)
    \\[.5ex]
    A &\coloneqq~ a \mid A\vlan A \mid A\vlor A \mid \exists x.A \mid \forall x.A
\end{array}
\end{equation*}
where the arity of $t$ and $p$ is $n$. Note, that in this paper we
consider the truth constants $\ttt$ (\emph{true}) and $\fff$
(\emph{false}) as atoms, and we consider all formulas in negation
normal form. The \bfit{negation} $(\dual\cdot)$ is defined for all
atoms and formulas via the De Morgan laws as follows:
\begin{equation*}
  \begin{array}{cc}
  \dual\ttt =\fff &
  \dual{p(t_1,\ldots,t_n)} = \dual{p}(t_1,\ldots,t_n) \\[.5ex]
  \dual\fff = \ttt &
  \dual{\dual p(t_1,\ldots,t_n)} = p(t_1,\ldots,t_n) \\[.5ex]
  \dual{A \vlan B} = \dual{A} \ \vlor \ \dual{B} &
  \dual{\exists x.A} = \forall x.\dual{A} \\[.5ex]
  \dual{A \vlor B} = \dual{A} \ \vlan \ \dual{B} & 
  \dual{\forall x.A} = \exists x.\dual{A} 
  \end{array}
\end{equation*}

@@

 A formula is
\bfit{rectified} if all bound variables are distinct from one another and from
all free variables. As every formula has a logically equivalent
rectified form, we will restrict most of the results in this paper to rectified formulas.
